---
title: "Reproducible Research Peer Assessment One"
author: "Bart Browning"
date: "11/10/2014"
output:
  html_document:
    keep_md: yes
---

##1. Read and Process the dataset:##
```{r read-proc-dataset}
steps  <- read.csv(file="../../activity.csv",head=TRUE,sep=",", colClasses=c(NA,"Date",NA))
notNa <- subset(steps, steps!="NA")
notNaStepsPerDay <- aggregate(steps ~ date, data = notNa, FUN = sum)
```

##2. Mean total number of steps taken per day:##

+ Histogram of total number of steps taken per day:

```{r hist-total-steps-per-day}
hist(notNaStepsPerDay$steps, main="", xlab="steps per day (NA values removed)")
```

+ the __mean__ number of steps taken each day is: `r sprintf("%.2f", mean(notNaStepsPerDay$steps))`
+ the __median__ number of steps taken each day is: `r sprintf("%.2f", median(notNaStepsPerDay$steps))`

##3. Steps taken (averaged across all days) versus the 5-minute intervals:##
```{r steps-taken-during-each-interval}
notNaAverageStepsPerInterval <- aggregate(steps ~ interval, data = notNa, FUN = mean)
plot(notNaAverageStepsPerInterval, type="l")  
maxInterval <- notNaAverageStepsPerInterval[which.max(notNaAverageStepsPerInterval$steps),]$interval
```

+ The 5-minute-interval that on average contains the maximum number of steps is: `r maxInterval`

##4. Imputed-data strategy: replace all NA values in the original data with the average-value for the relevant 5-minute-interval:##
```{r imputed-data}
isNa     <- subset(steps, is.na(steps))[,c(2,3)]                                                                                            
imputeNa <- merge(isNa, notNaAverageStepsPerInterval, by.x="interval", by.y="interval")                                               
imputed  <- rbind(imputeNa, notNa)                                                                                                     
```

+ Re-compute the histogram from step #2, but with imputed data added in place of NA:
```{r re-compute-histogram}
 imputedStepsPerDay <- aggregate(steps ~ date, data = imputed, FUN = sum)                                                               
 hist(imputedStepsPerDay$steps, main="", xlab="steps per day (NA values replaced)")                                                                                                         
```

+ the __mean__ number of steps taken each day is now: `r sprintf("%.2f", mean(imputedStepsPerDay$steps))`
+ Difference from step #2: `r sprintf("%.2f", mean(imputedStepsPerDay$steps) - mean(notNaStepsPerDay$steps))`
+ the __median__ number of steps taken each day is now: `r sprintf("%.2f", median(imputedStepsPerDay$steps))`
+ Difference from step #2: `r sprintf("%.2f", median(imputedStepsPerDay$steps) - median(notNaStepsPerDay$steps))`

##5. Average steps taken per 5-minute-interval (weekdays versus weekends):##
``` {r weekday-versus-weekend, fig.height=8}
library(timeDate)                                                                                                                     
newCol      <- ifelse(isWeekday(imputed$date), "weekday", "weekend")                                                                       
imputedWday <- cbind(imputed, newCol)                                                                                                   
weekday     <- subset(imputedWday, newCol=="weekday")                                                                                     
weekend     <- subset(imputedWday, newCol=="weekend")                                                                                     
weekdayASPI <- aggregate(steps ~ interval, data = weekday, FUN = mean)                                                                
weekendASPI <- aggregate(steps ~ interval, data = weekend, FUN = mean)                                                                
par(mfrow=c(2,1))                                                                                                                     
plot(weekdayASPI, type="l", main="Weekday avg steps per interval")
plot(weekendASPI, type="l", main="Weekend avg steps per interval")
```

